---
title: "MRI Attachment - Regression analyses for Amy/Hip structure and attachment behaviours (transformed behavioural data)"
author: "LJS"
date: "2023-02-28"
output: html_document
fig_height: 10
---

# Libraries 

```{r setup, include=FALSE}
library("rlang") #toolbox
library("vctrs") #data manipulation
library("dplyr") #data manipulation
library("ggplot2") #plotting
library("scales") #plotting
library("here") #toolbox
library("tidyverse") #data manipulation
library("corrplot") #plotting
library("ggpubr") #plotting
library("tidytext") #data manipulation
library("rstatix") #statistical tests
library("OpenMx") #residualise metrics
library("umx") #residualise metrics
library("knitr") #rmarkdown
library("htmltools") #rmarkdown
library("markdown") #rmarkdown
library("httpuv") #rmarkdown
library("NCmisc") #toolbox
library("performance") #check models performance
library("easystats") #check models performance
library("see") #check models performance
library("car") #logit transformation
library("lmtest") #check models performance
knitr::opts_chunk$set(warning = FALSE, message = FALSE) #do not print warning messages
```

# Neonatal MRI and infant attachment

## Section I: Prepare dataframe with data and transform potential covariates

Start from spreadsheet with raw times from coding scheme (Williams and Turner, 2020). 
IDs are already filtered in this spreadhseet, based on MRI + attachment criteria. 
Merge attachment, MRI and covariates data in a dataframe.

### Create spreadsheet for all statistical analyses

```{r}
Attachment <- read.csv(here("raw_data", "IDs_WT_rawtimes_filtered.csv"), header=T, dec=".",sep=",")
MRI_structural <- read.csv(here("raw_data", "IDs_MRI_structural.csv"), header=T, dec=".",sep=",")
MRI_DTI <- read.csv(here("raw_data", "IDs_MRI_DTI_gmNODDI.csv"), header=T, dec=".",sep=",")
Covariates <- read.csv(here("raw_data", "IDs_covariates.csv"), header=T, dec=".",sep=",")
Data <- merge(Attachment, MRI_DTI, by="ID")
Data <- merge(Data, MRI_structural, by="ID")
Data <- merge(Data, Covariates, by="ID")
rownames(Data)<-Data$ID
Data$ID <- NULL
Data$Ethnicity <- NULL
```

Avoid scientific notation.

```{r}
options(scipen = 999)
```

Transform variables to categorical nominal (factors without order) or categorical ordinal (factors with order) accordingly.

```{r}
Data$sex = as.factor(Data$sex)
Data$Maternal_smoking = as.factor(Data$Maternal_smoking)
Data$Maternal_alcohol = as.factor(Data$Maternal_alcohol)
Data$Maternal_education = factor(Data$Maternal_education, order = TRUE, levels =c("1", "2", "3", "4", "5", "6", "7"))
Data$Maternal_antenatal_anxiety = as.factor(Data$Maternal_antenatal_anxiety)
Data$Maternal_antenatal_depression = as.factor(Data$Maternal_antenatal_depression)
```

Scale brain data and residualise brain against age at scan (these keep their original name for simplicity).

```{r}
Data[,c(4:24)] <- scale(Data[,c(4:24)])
Data <- umx_residualize(c("vol_hip_l", "vol_hip_r", "vol_amy_l", "vol_amy_r", "MD_hip_l", "MD_hip_r", "MD_amy_l", "MD_amy_r", "FA_hip_l", "FA_hip_r", "FA_amy_l", "FA_amy_r", "NDI_hip_l", "NDI_hip_r", "NDI_amy_l", "NDI_amy_r", "ODI_hip_l", "ODI_hip_r", "ODI_amy_l", "ODI_amy_r"), cov = c("age_scan"), data = Data)
```

Transform HD, FNF and AA raw scores

Logit transformation and scale variables.

```{r}
Data$FNF_Raw = logit(Data$FNF_Raw)
Data[,c(1:3)] <- scale(Data[,c(1:3)])
```


## Associations between MRI features and attachment behaviours

Run models for brain data as independent variables, and HD Raw score as dependent variable.
Set variables for loops.

```{r}
targetdata=Data
colns_oi = colnames(Data)[grep('_hip|_amy',colnames(targetdata))]
```

### Brain data ~ HD Raw score

Run models and create table with results.

```{r}
for(i in colns_oi){
  formula.torun = paste0('HD_Raw~Maternal_education+Maternal_antenatal_anxiety+', i)
  fit=lm(as.formula(formula.torun),data=targetdata) 
  report=summary(fit)
  single.model=report$coefficients[nrow(report$coefficients),]
  
  if (i==colns_oi[1]){
    total.report=single.model
  }else{
    total.report=rbind(total.report,single.model)
  }
  
}

rownames(total.report)=colns_oi
Total.report.HD = data.frame(total.report)
colnames(Total.report.HD) <- c("Beta", "SE", "t-value", "p-value")
Total.report.HD$Brain <- rownames(Total.report.HD)
rownames(Total.report.HD) <- NULL
Total.report.HD$Behaviour = "HD_Raw"
new_order = c("Brain","Behaviour","Beta", "SE", "t-value", "p-value")
Total.report.HD <- Total.report.HD[, new_order]
Results.report <- Total.report.HD
```

#### Model diagnostics for Brain data ~ HD Raw score

```{r, fig.fullheight=TRUE}
model_list=list()
for(i in colns_oi){
  formula.torun = paste0('HD_Raw~Maternal_education+Maternal_antenatal_anxiety+', i)
  fit=lm(as.formula(formula.torun),data=targetdata)
  model_list[[i]]=fit}

lapply(model_list, check_model)

#Double-check normality of residuals with Q-Q plots
for (i in model_list) {
  plot(i, which=2)}
```

According to above, models do not show issues with linearity of the association, normality of residuals, homogeneity of variance or multicollinearity of predictor variables. 


### Brain data ~ FNF Raw score

Run models and create table with results.

```{r}
for(i in colns_oi){
  formula.torun = paste0('FNF_Raw~Maternal_education++Maternal_antenatal_anxiety+', i)
  fit=lm(as.formula(formula.torun),data=targetdata) 
  report=summary(fit)
  single.model=report$coefficients[nrow(report$coefficients),]
  
  if (i==colns_oi[1]){
    total.report=single.model
  }else{
    total.report=rbind(total.report,single.model)
  }
  
}

rownames(total.report)=colns_oi
Total.report.FNF = data.frame(total.report)
colnames(Total.report.FNF) <- c("Beta", "SE", "t-value", "p-value")
Total.report.FNF$Brain <- rownames(Total.report.FNF)
rownames(Total.report.FNF) <- NULL
Total.report.FNF$Behaviour = "FNF_Raw"
new_order = c("Brain","Behaviour","Beta", "SE", "t-value", "p-value")
Total.report.FNF <- Total.report.FNF[, new_order]
Results.report <- rbind (Results.report, Total.report.FNF)
```

#### Model diagnostics for Brain data ~ FNF Raw score

```{r, fig.fullheight=TRUE}
model_list=list()
for(i in colns_oi){
  formula.torun = paste0('FNF_Raw~Maternal_education+Maternal_antenatal_anxiety+', i)
  fit=lm(as.formula(formula.torun),data=targetdata)
  model_list[[i]]=fit}

lapply(model_list, check_model)
lapply(model_list, plot)

#Double-check normality of residuals with Q-Q plots
for (i in model_list) {
  plot(i, which=2)}
```

##### Check homogeneity of variance

```{r}
lapply(model_list, gqtest)
```

According to above, models show improvements with linearity of the association, normality of residuals, and homogeneity of variance. 


### Brain data ~ AA Raw score

Run models and create table with results.

```{r}
for(i in colns_oi){
  formula.torun = paste0('AA_Raw~Maternal_education+Maternal_antenatal_anxiety+', i)
  fit=lm(as.formula(formula.torun),data=targetdata) 
  report=summary(fit)
  single.model=report$coefficients[nrow(report$coefficients),]
  
  if (i==colns_oi[1]){
    total.report=single.model
  }else{
    total.report=rbind(total.report,single.model)
  }
  
}

rownames(total.report)=colns_oi
Total.report.AA = data.frame(total.report)
colnames(Total.report.AA) <- c("Beta", "SE", "t-value", "p-value")
Total.report.AA$Brain <- rownames(Total.report.AA)
rownames(Total.report.AA) <- NULL
Total.report.AA$Behaviour = "AA_Raw"
new_order = c("Brain","Behaviour","Beta", "SE", "t-value", "p-value")
Total.report.AA <- Total.report.AA[, new_order]
Results.report <- rbind (Results.report, Total.report.AA)
```

#### Model diagnostics for Brain data ~ AA Raw score

```{r, fig.fullheight=TRUE}
model_list=list()
for(i in colns_oi){
  formula.torun = paste0('AA_Raw~Maternal_education+Maternal_antenatal_anxiety+', i)
  fit=lm(as.formula(formula.torun),data=targetdata)
  model_list[[i]]=fit}

lapply(model_list, check_model)

#Double-check normality of residuals with Q-Q plots
for (i in model_list) {
  plot(i, which=2)}
```

According to above, models do not show issues with linearity of the association, normality of residuals, homogeneity of variance or multicollinearity of predictor variables. 


### create total reports and adjust p-values

```{r}
Results.report.hip.l <- filter(Results.report,grepl("hip_l",Brain))
Results.report.hip.l$"p-corrected"=p.adjust(Results.report.hip.l[,'p-value'],method='BH')

Results.report.hip.r <- filter(Results.report,grepl("hip_r",Brain))
Results.report.hip.r$"p-corrected"=p.adjust(Results.report.hip.r[,'p-value'],method='BH')

Results.report.amy.l <- filter(Results.report,grepl("amy_l",Brain))
Results.report.amy.l$"p-corrected"=p.adjust(Results.report.amy.l[,'p-value'],method='BH')

Results.report.amy.r <- filter(Results.report,grepl("amy_r",Brain))
Results.report.amy.r$"p-corrected"=p.adjust(Results.report.amy.r[,'p-value'],method='BH')

Total.report = rbind(Results.report.hip.l, Results.report.hip.r, Results.report.amy.l, Results.report.amy.r)
Total.report$siglabel <- NA
Total.report$siglabel[Total.report$"p-corrected"<0.05] = "*"
Total.report$siglabel[Total.report$"p-corrected"<0.01] = "**"
Total.report$siglabel[Total.report$"p-corrected"<0.005] = "***"
  
write.csv(Total.report, here("results/regression", "GM.reg.transformed.csv"), row.names = FALSE)
```

## Plot associations between MRI features and attachment behaviours

```{r, fig.fullheight=TRUE}
data <- read.csv(here("results/regression", "GM.reg.csv"), header=T, dec=".",sep=",")

data$Brain <- factor(data$Brain, levels=c("vol_amy_l", "FA_amy_l", "MD_amy_l", "NDI_amy_l", "ODI_amy_l",
                                  "vol_amy_r", "FA_amy_r", "MD_amy_r", "NDI_amy_r", "ODI_amy_r",
                                  "vol_hip_l", "FA_hip_l", "MD_hip_l", "NDI_hip_l", "ODI_hip_l",
                                  "vol_hip_r", "FA_hip_r", "MD_hip_r", "NDI_hip_r", "ODI_hip_r", 
                                  "HD_Raw", "FNF_Raw", "AA_Raw"))
g <- ggplot(data = data, mapping = aes(x=Brain, y = Behaviour, fill = Beta)) +
  scale_fill_gradient2(mid="#FBFEF9",low="#841026",high="#0E6885", limits=c(-0.5,0.5)) +
  geom_tile() +
  scale_y_reordered() +
  ylab(label = "Attachment dimension") 

g <- g + geom_text(aes(label=siglabel), vjust=0.6, hjust=0.5) + theme_bw()

g

ggsave(here("results/figures", "GM.reg.heatmap.transformed.pdf"), plot = g, dpi=300)
```


```{r}
markdownToHTML("MRI_Attachment_GM_reg_transformed.Rmd",output ="MRI_Attachment_GM_reg_transformed.html")
```

