---
title: "Interater reliability analyses"
author: "LJS"
date: "2024-03-11"
output: html_document
---

# Libraries 

```{r}
library("irr") #reliability analyses
library("knitr") #rmarkdown
library("htmltools") #rmarkdown
library("markdown") #rmarkdown
library("httpuv") #rmarkdown
```

# Neonatal MRI and infant attachment

## Section I: Prepare dataframe with data

Start from spreadsheet with raw scores and attachment styles (Williams and Turner, 2020) from first and second coder. 

### Create spreadsheet

```{r}
Data <- read.csv(here("raw_data", "IDs_WT_inter.csv"), header=T, dec=".",sep=",")
```

## Create table to add results

```{r}
results_table <- data.frame(matrix(ncol = 6, nrow = 4))
x <- c("measure", "ICC/Kappa", "lbound", "ubound", "%agreement", "pval")
colnames(results_table) <- x
results_table$measure <- c("HD_Raw", "FNF_Raw", "AA_Raw", "AttachmentStyle")
```

## Section II: Calculate ICC for raw scores

### HD Raw scores

```{r}

# data wrangling

Data_HD <- Data[,c(1:3)]
Data_HD_1 <- Data_HD[Data_HD$Rater=="1",]
Data_HD_1$Rater <- NULL
colnames(Data_HD_1) <- c("ID", "1")
Data_HD_2 <- Data_HD[Data_HD$Rater=="2",]
Data_HD_2$Rater <- NULL
colnames(Data_HD_2) <- c("ID", "2")

Data_HD <- merge(Data_HD_1, Data_HD_2, by="ID")

rownames(Data_HD) <- Data_HD$ID
Data_HD$ID <- NULL

# calculate ICC

ICC_HD <- icc(
  Data_HD, model = "twoway", 
  type = "agreement", unit = "single"
)

# write into the results table

results_table$"ICC/Kappa"[which(results_table$measure=="HD_Raw")] <- ICC_HD$value
results_table$lbound[which(results_table$measure=="HD_Raw")] <- ICC_HD$lbound
results_table$ubound[which(results_table$measure=="HD_Raw")] <- ICC_HD$ubound
results_table$pval[which(results_table$measure=="HD_Raw")] <- ICC_HD$p.value

```

### FNF Raw scores

```{r}

# data wrangling

Data_FNF <- Data[,c(1,2,4)]
Data_FNF_1 <- Data_FNF[Data_FNF$Rater=="1",]
Data_FNF_1$Rater <- NULL
colnames(Data_FNF_1) <- c("ID", "1")
Data_FNF_2 <- Data_FNF[Data_FNF$Rater=="2",]
Data_FNF_2$Rater <- NULL
colnames(Data_FNF_2) <- c("ID", "2")

Data_FNF <- merge(Data_FNF_1, Data_FNF_2, by="ID")

rownames(Data_FNF) <- Data_FNF$ID
Data_FNF$ID <- NULL

# calculate ICC

ICC_FNF <- icc(
  Data_FNF, model = "twoway", 
  type = "agreement", unit = "single"
)

# write into the results table

results_table$"ICC/Kappa"[which(results_table$measure=="FNF_Raw")] <- ICC_FNF$value
results_table$lbound[which(results_table$measure=="FNF_Raw")] <- ICC_FNF$lbound
results_table$ubound[which(results_table$measure=="FNF_Raw")] <- ICC_FNF$ubound
results_table$pval[which(results_table$measure=="FNF_Raw")] <- ICC_FNF$p.value

```

### AA Raw scores

```{r}

# data wrangling

Data_AA <- Data[,c(1,2,5)]
Data_AA_1 <- Data_AA[Data_AA$Rater=="1",]
Data_AA_1$Rater <- NULL
colnames(Data_AA_1) <- c("ID", "1")
Data_AA_2 <- Data_AA[Data_AA$Rater=="2",]
Data_AA_2$Rater <- NULL
colnames(Data_AA_2) <- c("ID", "2")

Data_AA <- merge(Data_AA_1, Data_AA_2, by="ID")

rownames(Data_AA) <- Data_AA$ID
Data_AA$ID <- NULL

# calculate ICC

ICC_AA <- icc(
  Data_AA, model = "twoway", 
  type = "agreement", unit = "single"
)

# write into the results table

results_table$"ICC/Kappa"[which(results_table$measure=="AA_Raw")] <- ICC_AA$value
results_table$lbound[which(results_table$measure=="AA_Raw")] <- ICC_AA$lbound
results_table$ubound[which(results_table$measure=="AA_Raw")] <- ICC_AA$ubound
results_table$pval[which(results_table$measure=="AA_Raw")] <- ICC_AA$p.value

```

## Section III: Calculate Kappa for attachment styles

```{r}

# data wrangling

Data_AttachmentStyles <- Data[,c(1,2,6)]
Data_AttachmentStyles_1 <- Data_AttachmentStyles[Data_AttachmentStyles$Rater=="1",]
Data_AttachmentStyles_1$Rater <- NULL
colnames(Data_AttachmentStyles_1) <- c("ID", "1")
Data_AttachmentStyles_2 <- Data_AttachmentStyles[Data_AttachmentStyles$Rater=="2",]
Data_AttachmentStyles_2$Rater <- NULL
colnames(Data_AttachmentStyles_2) <- c("ID", "2")

Data_AttachmentStyles <- merge(Data_AttachmentStyles_1, Data_AttachmentStyles_2, by="ID")

rownames(Data_AttachmentStyles) <- Data_AttachmentStyles$ID
Data_AttachmentStyles$ID <- NULL

# calculate percentage of agreement

Agree_AttachmentStyles <- agree(Data_AttachmentStyles, tolerance=0)

# calculate kappa

Kappa_AttachmentStyles <- kappa2(Data_AttachmentStyles)

# write into the results table

results_table$"ICC/Kappa"[which(results_table$measure=="AttachmentStyle")] <- Kappa_AttachmentStyles$value
results_table$"%agreement"[which(results_table$measure=="AttachmentStyle")] <- Agree_AttachmentStyles$value
results_table$pval[which(results_table$measure=="AttachmentStyle")] <- Kappa_AttachmentStyles$p.value

```

## Section IV: save results

```{r}
write.csv(results_table, here("results/reliability", "WT_inter_rel.csv"), row.names = FALSE)
```

```{r}
markdownToHTML("MRI_Attachment_inter_rel.Rmd",output ="MRI_Attachment_inter_rel.html")
```

