---
title: "MRI Attachment - Covariate analyses for Amy/Hip structure and attachment behaviours"
author: "LJS"
date: "2023-02-28"
output: html_document
---

# Libraries 

```{r setup, include=FALSE}
library("rlang") #toolbox
library("dplyr") #data manipulation
library("ggplot2") #plotting
library("scales") #plotting
library("here") #toolbox
library("tidyverse") #data manipulation
library("ggpubr") #plotting
library("tidytext") #data manipulation
library("OpenMx") #residualise metrics
library("umx") #residualise metrics
library("knitr") #rmarkdown
library("htmltools") #rmarkdown
library("markdown") #rmarkdown
library("httpuv") #rmarkdown
library("vtable") #summary tables
library("gtsummary") #summary tables
library("NCmisc") #toolbox
library("Hmisc") #statistical tests
library("stringr") #data manipulation
knitr::opts_chunk$set(warning = FALSE, message = FALSE) #do not print warning messages
```

# Neonatal MRI and infant attachment

## Section I: Prepare dataframe with data and transform potential covariates

Start from spreadsheet with raw times from coding scheme (Williams and Turner, 2020). 
IDs are already filtered in this spreadhseet, based on MRI + attachment criteria. 
Merge attachment, MRI and covariates data in a dataframe.

### Create spreadsheet for all analyses

```{r}
Attachment <- read.csv(here("raw_data", "IDs_WT_rawtimes_filtered.csv"), header=T, dec=".",sep=",")
MRI_structural <- read.csv(here("raw_data", "IDs_MRI_structural.csv"), header=T, dec=".",sep=",")
MRI_DTI <- read.csv(here("raw_data", "IDs_MRI_DTI_gmNODDI.csv"), header=T, dec=".",sep=",")
Covariates <- read.csv(here("raw_data", "IDs_covariates.csv"), header=T, dec=".",sep=",")
Data <- merge(Attachment, MRI_DTI, by="ID")
Data <- merge(Data, MRI_structural, by="ID")
Data <- merge(Data, Covariates, by="ID")
rownames(Data)<-Data$ID
Data$ID <- NULL
```

Avoid scientific notation.

```{r}
options(scipen = 999)
```

### Transform potential covariates to categorical nominal (factors without order) or categorical ordinal (factors with order) accordingly.

```{r}
Data$sex = as.factor(Data$sex)
Data$Maternal_smoking = as.factor(Data$Maternal_smoking)
Data$Maternal_alcohol = as.factor(Data$Maternal_alcohol)
Data$Maternal_education = factor(Data$Maternal_education, order = TRUE, levels =c("1", "2", "3", "4", "5", "6", "7"))
Data$Maternal_antenatal_anxiety = as.factor(Data$Maternal_antenatal_anxiety)
Data$Maternal_antenatal_depression = as.factor(Data$Maternal_antenatal_depression)
```

## Get residualised metrics

To control for head size, MRI metrics need to be residualised against age at scan or total brain volume. 

First, the colinearity of age at scan and total tissue volume is checked.

```{r}
shapiro.test(Data$age_scan)
hist(Data$age_scan)
shapiro.test(Data$vol_t)
hist(Data$vol_t)
```

According to above, both measures follow a normal distribution so Pearson correlation is chosen.

```{r}
cor(Data$age_scan, Data$vol_t, method="pearson")
```

Both measures are highly correlated so one of them will be used to control for head size. 

Age at scan would be more suitable to adjust NODDI metrics, so MRI metrics are residualised against age at scan before starting covariate analyses.

Residualise MRI metrics against age at scan (these keep their original name for simplicity).

```{r}
Data$Ethnicity <- NULL
Data[,c(1:24)] <- scale(Data[,c(1:24)])

Data <- umx_residualize(c("vol_hip_l", "vol_hip_r", "vol_amy_l", "vol_amy_r", "MD_hip_l", "MD_hip_r", "MD_amy_l", "MD_amy_r", "FA_hip_l", "FA_hip_r", "FA_amy_l", "FA_amy_r", "NDI_hip_l", "NDI_hip_r", "NDI_amy_l", "NDI_amy_r", "ODI_hip_l", "ODI_hip_r", "ODI_amy_l", "ODI_amy_r"), cov = c("age_scan"), data = Data)
```


## Covariate analyses

### Categorical variables

Categorical variables: Sex, maternal smoking, maternal alcohol, maternal antenatal depression, maternal antenatal anxiety.
Check normality of continuous variables (MRI and behavioural variables).

Shapiro tests.

```{r}
targetdata=Data[,grep('HD_Raw|FNF_Raw|AA_Raw|FA_|MD_|NDI_|ODI_|vol_', colnames(Data))]
targetdata=as.data.frame(targetdata)
apply(targetdata,2,shapiro.test)
```

According to above: all behavioural variables (HD_Raw, FNF_Raw, and AA_Raw) and some MRI variables (MD_hip_l, MD_hip_r, MD_amy_l, MD_amy_r, NDI_amy_l, NDI_hip_r, NDI_amy_l, ODI_hip_l, ODI_hip_r, ODI_amy_l, vol_hip_r, vol_amy_l) seem non-normally distributed.
Use histograms to double-check distribution of these variables visually.

```{r}
targetdata=Data[,grep("HD_Raw|FNF_Raw|AA_Raw|MD_hip_l|MD_hip_r|MD_amy_l|MD_amy_r|NDI_hip_l|NDI_hip_r|NDI_amy_l|ODI_hip_l|ODI_hip_r|ODI_amy_l|vol_hip_r|vol_amy_l", colnames(Data))]
targetdata=as.data.frame(targetdata)
for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% hist(main = cols)
}
```

In case of doubt, normal QQ plots can be used to check distribution of variables visually.

```{r}
for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% qqnorm(main = cols)
}
```

According to above: FNF_Raw, MD_hip_l, MD_hip_r, MD_amy_r, NDI_hip_r, ODI_hip_r, vol_hip_r seem non-normally distributed.

Display tables of descriptive statistics and tests by confounder.

```{r}
Pval_fun <- function(x) style_pvalue(x, digits = 3)
Stat_fun <- function(x) style_sigfig(x, digits=3)
```

Associations between sex and MRI/behavioural variables.

```{r}
targetdata <- Data[,c(1:20,22,23,24,25)]

table_cov <- targetdata %>% tbl_summary(by = sex,
  label = NULL,
  statistic = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ c("{median} ({min}, {max})"), 
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "MD_amy_l", "FA_hip_l", "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") ~ c("{mean} ({sd})"), all_categorical() ~ "{n} ({p}%)"),
  digits = everything()~3) %>% 
  add_p(
  test = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ "wilcox.test",
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "FA_hip_l","MD_amy_l",  "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") 
              ~ "t.test", all_categorical() ~ "fisher.test"),
  pvalue_fun = Pval_fun) %>% 
  separate_p_footnotes() %>% 
  modify_header(
    list(
      statistic ~ "**t-statistic**",
      p.value ~ "**p-value**"
    )
  ) %>%
  modify_fmt_fun(
    list(
      statistic ~ Stat_fun,
      p.value ~ Pval_fun
    ))

table_cov
table_cov <- as_tibble(table_cov)
colnames(table_cov) <- c("Characteristic", "M", "F", "t-statistic", "p-value")

write.csv(table_cov, here("results/covariates", "GM.Cov.sex.table.csv"), row.names=FALSE)
```

Associations between maternal smoking and MRI/behavioural variables.

```{r}
targetdata <- Data[,c(1:20,22,23,24,33)]

table_cov <- targetdata %>% tbl_summary(by = Maternal_smoking,
  label = NULL,
  statistic = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ c("{median} ({min}, {max})"), 
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "MD_amy_l", "FA_hip_l", "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") ~ c("{mean} ({sd})"), all_categorical() ~ "{n} ({p}%)"),
  digits = everything()~3) %>% 
  add_p(
  test = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ "wilcox.test",
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "FA_hip_l","MD_amy_l",  "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") 
              ~ "t.test", all_categorical() ~ "fisher.test"),
  pvalue_fun = Pval_fun) %>% 
  separate_p_footnotes() %>% 
  modify_header(
    list(
      statistic ~ "**t-statistic**",
      p.value ~ "**p-value**"
    )
  ) %>%
  modify_fmt_fun(
    list(
      statistic ~ Stat_fun,
      p.value ~ Pval_fun
    ))

table_cov
table_cov <- as_tibble(table_cov)
colnames(table_cov) <- c("Characteristic", "None", "Any", "t-statistic", "p-value")
write.csv(table_cov, here("results/covariates", "GM.Cov.smo.table.csv"), row.names = FALSE)
```

Associations between maternal alcohol and MRI/behavioural variables.

```{r}
targetdata <- Data[,c(1:20,22,23,24,34)]

table_cov <- targetdata %>% tbl_summary(by = Maternal_alcohol,
  label = NULL,
  statistic = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ c("{median} ({min}, {max})"), 
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "MD_amy_l", "FA_hip_l", "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") ~ c("{mean} ({sd})"), all_categorical() ~ "{n} ({p}%)"),
  digits = everything()~3) %>% 
  add_p(
  test = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ "wilcox.test",
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "FA_hip_l","MD_amy_l",  "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") 
              ~ "t.test", all_categorical() ~ "fisher.test"),
  pvalue_fun = Pval_fun) %>% 
  separate_p_footnotes() %>% 
  modify_header(
    list(
      statistic ~ "**t-statistic**",
      p.value ~ "**p-value**"
    )
  ) %>%
  modify_fmt_fun(
    list(
      statistic ~ Stat_fun,
      p.value ~ Pval_fun
    ))

table_cov
table_cov <- as_tibble(table_cov)
colnames(table_cov) <- c("Characteristic", "None", "Any", "t-statistic", "p-value")
write.csv(table_cov, here("results/covariates", "GM.Cov.alc.table.csv"), row.names = FALSE)

```

Associations between maternal antenatal depression and MRI/behavioural variables.

```{r}
targetdata <- Data[,c(1:20,22,23,24,35)]

table_cov <- targetdata %>% tbl_summary(by = Maternal_antenatal_depression,
  label = NULL,
  statistic = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ c("{median} ({min}, {max})"), 
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "MD_amy_l", "FA_hip_l", "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") ~ c("{mean} ({sd})"), all_categorical() ~ "{n} ({p}%)"),
  digits = everything()~3) %>% 
  add_p(
  test = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ "wilcox.test",
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "FA_hip_l","MD_amy_l",  "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") 
              ~ "t.test", all_categorical() ~ "fisher.test"),
  pvalue_fun = Pval_fun) %>% 
  separate_p_footnotes() %>% 
  modify_header(
    list(
      statistic ~ "**t-statistic**",
      p.value ~ "**p-value**"
    )
  ) %>%
  modify_fmt_fun(
    list(
      statistic ~ Stat_fun,
      p.value ~ Pval_fun
    ))

table_cov
table_cov <- as_tibble(table_cov)
colnames(table_cov) <- c("Characteristic", "Absence", "Presence", "t-statistic", "p-value")
write.csv(table_cov, here("results/covariates", "GM.Cov.dep.table.csv"), row.names = FALSE)
```

Associations between maternal antenatal anxiety and MRI/behavioural variables.

```{r}
targetdata <- Data[,c(1:20,22,23,24,36)]

table_cov <- targetdata %>% tbl_summary(by = Maternal_antenatal_anxiety,
  label = NULL,
  statistic = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ c("{median} ({min}, {max})"), 
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "MD_amy_l", "FA_hip_l", "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") ~ c("{mean} ({sd})"), all_categorical() ~ "{n} ({p}%)"),
  digits = everything()~3) %>% 
  add_p(
  test = list(c("FNF_Raw", "MD_hip_l", "MD_hip_r", "MD_amy_r", "NDI_hip_r", "vol_hip_r", "ODI_hip_r") ~ "wilcox.test",
                   c("HD_Raw", "AA_Raw", "vol_hip_l", "FA_hip_l","MD_amy_l",  "NDI_hip_l", "ODI_hip_l", "FA_hip_r","vol_amy_l", "FA_amy_l", "NDI_amy_l", "ODI_amy_l", "vol_amy_r", "FA_amy_r", "NDI_amy_r", "ODI_amy_r") 
              ~ "t.test", all_categorical() ~ "fisher.test"),
  pvalue_fun = Pval_fun) %>% 
  separate_p_footnotes() %>% 
  modify_header(
    list(
      statistic ~ "**t-statistic**",
      p.value ~ "**p-value**"
    )
  ) %>%
  modify_fmt_fun(
    list(
      statistic ~ Stat_fun,
      p.value ~ Pval_fun
    ))


table_cov
table_cov <- as_tibble(table_cov)
colnames(table_cov) <- c("Characteristic", "Absence", "Presence", "t-statistic", "p-value")
write.csv(table_cov, here("results/covariates", "GM.Cov.anx.table.csv"), row.names = FALSE)

```

According to above, maternal antenatal anxiety meets the definition of confounder and was thereby selected as covariate in further analyses.

### Continuous/categorical ordinal variables 

Continuous variables: GA at birth, SIMD, maternal age.
Shapiro tests.

```{r}
targetdata=Data[,grep('GA_birth|SIMD|Maternal_age', colnames(Data))]
targetdata=as.data.frame(targetdata)
apply(targetdata,2,shapiro.test)
```

According to above: GA at birth, and SIMD seem non-normally distributed. Maternal age is normally distributed.
Use histograms to double-check distribution of these variables visually.

```{r}
for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% hist(main = cols)
}
```

In case of doubt, normal QQ plots can be used to check distribution of variables visually.

```{r}
for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% qqnorm(main = cols)
}
```

According to above, GA at birth, and SIMD seem non-normally distributed. 
Run correlations (Spearman tests for Maternal education, GA at birth and SIMD; Spearman/Pearson tests for Maternal age, depending on the distribution of the MRI/behavioural variable).

```{r, out.width="50%"}
targetdata <- Data[,c(1:20,22,23,24,27,28,32)]

cors <- function(df) { 
  # turn all three matrices (r, n, and P into a data frame)
  M <- Hmisc::rcorr(as.matrix(df), type="spearman")
  # return the three data frames in a list return(Mdf)
  Mdf <- map(M, ~data.frame(.x))
}

newdata <- cors(targetdata)

formatted_cors <- function(df){
  cors(df) %>%
    map(~rownames_to_column(.x, var="measure1")) %>%
    map(~pivot_longer(.x, -measure1, names_to="measure2")) %>% 
    bind_rows(.id = "id") %>%
    tidyr::pivot_wider(names_from = id, values_from = value) %>%
    mutate(sig_label = case_when(
      P<0.05 & P>0.01 ~ "*", 
      P<0.01 & P>0.001 ~ "**", 
      P<0.001 ~ "***",
      TRUE ~ " "))}

data_spearman <- formatted_cors(targetdata)

data_spearman = as.data.frame(data_spearman)

colnames(data_spearman) = c("Y", "X", "Estimate", "N", "P_value", "sig_label")

data_spearman <- data_spearman %>% filter(Y == "Maternal_education" | Y == "SIMD" |  Y == "GA_birth" )

data_spearman <- data_spearman %>% filter(
  X == "vol_hip_l" | X == "FA_hip_l" |  X == "MD_hip_l" | X == "NDI_hip_l" | X == "ODI_hip_l" |  
    X == "vol_hip_r" |   X == "FA_hip_r" |  X == "MD_hip_r" | X == "NDI_hip_r" | X == "ODI_hip_r" |
    X == "vol_amy_l" | X == "FA_amy_l" | X == "MD_amy_l" |  X == "NDI_amy_l" |   X == "ODI_amy_l" |
    X == "vol_amy_r" | X == "FA_amy_r" | X == "MD_amy_r" |  X == "NDI_amy_r" | X == "ODI_amy_r" | 
    X == "HD_Raw" |  X == "FNF_Raw"|  X == "AA_Raw" )


targetdata=Data[,grep("FNF_Raw|MD_hip_l|MD_hip_r|MD_amy_r|NDI_hip_r|vol_hip_r|ODI_hip_r|Maternal_age", colnames(Data))]

cors <- function(df) { 
  # turn all three matrices (r, n, and P into a data frame)
  M <- Hmisc::rcorr(as.matrix(df), type="spearman")
  # return the three data frames in a list return(Mdf)
  Mdf <- map(M, ~data.frame(.x))
}

formatted_cors <- function(df){
  cors(df) %>%
    map(~rownames_to_column(.x, var="measure1")) %>%
    map(~pivot_longer(.x, -measure1, names_to="measure2")) %>% 
    bind_rows(.id = "id") %>%
    pivot_wider(names_from = id, values_from = value) %>%
    mutate(sig_label = case_when(
      P<0.05 & P>0.01 ~ "*", 
      P<0.01 & P>0.001 ~ "**", 
      P<0.001 ~ "***",
      TRUE ~ " "))}

MA_spearman <- formatted_cors(targetdata)

MA_spearman = as.data.frame(MA_spearman)

colnames(MA_spearman) = c("Y", "X", "Estimate", "N", "P_value", "sig_label")

MA_spearman <- MA_spearman %>% filter(Y == "Maternal_age")

MA_spearman <- MA_spearman %>% filter(
    X == "MD_hip_l" |  
    X == "vol_hip_r" | X == "MD_hip_r" | X == "NDI_hip_r" | X == "ODI_hip_r" |
    X == "MD_amy_r" | 
    X == "FNF_Raw" )

data <- rbind(data_spearman, MA_spearman)


targetdata <- Data[,c(1:24,31)]
targetdata=Data[,-grep("FNF_Raw|MD_hip_l|MD_hip_r|MD_amy_r|NDI_hip_r|vol_hip_r|ODI_hip_r|Maternal_age", colnames(targetdata))]

cors <- function(df) { 
  # turn all three matrices (r, n, and P into a data frame)
  M <- Hmisc::rcorr(as.matrix(df), type="pearson")
  # return the three data frames in a list return(Mdf)
  Mdf <- map(M, ~data.frame(.x))
}

formatted_cors <- function(df){
  cors(df) %>%
    map(~rownames_to_column(.x, var="measure1")) %>%
    map(~pivot_longer(.x, -measure1, names_to="measure2")) %>% 
    bind_rows(.id = "id") %>%
    pivot_wider(names_from = id, values_from = value) %>%
    mutate(sig_label = case_when(
      P<0.05 & P>0.01 ~ "*", 
      P<0.01 & P>0.001 ~ "**", 
      P<0.001 ~ "***",
      TRUE ~ " "))}

MA_pearson <- formatted_cors(targetdata)

MA_pearson = as.data.frame(MA_pearson)

colnames(MA_pearson) = c("Y", "X", "Estimate", "N", "P_value", "sig_label")

MA_pearson <- MA_pearson %>% filter(Y == "Maternal_age")

MA_pearson <- MA_pearson %>% filter(
  X == "vol_hip_l" | X == "FA_hip_l" | X == "NDI_hip_l" | X == "ODI_hip_l" |  
    X == "FA_hip_r" |
    X == "vol_amy_l" | X == "FA_amy_l" | X == "MD_amy_l" |  X == "NDI_amy_l" |   X == "ODI_amy_l" |
    X == "vol_amy_r" | X == "FA_amy_r" |  X == "NDI_amy_r" | X == "ODI_amy_r" | 
    X == "HD_Raw" |  X == "AA_Raw" )

data <- rbind(data, MA_pearson)

kable(data, col.names = gsub("[.]", " ", names(data)))
write.csv(data, here("results/covariates", "GM.Cov.age.ed.GA.SIMD.table.csv"), row.names = FALSE)
```

According to above, Maternal education met the definition of confounder and was thereby used as a covariate in further analyses.

```{r}
markdownToHTML("MRI_Attachment_GM_cov.Rmd",output ="MRI_Attachment_GM_cov.html")
```


