---
title: "MRI Attachment - Demographics"
author: "LJS"
date: "2023-02-28"
output: html_document
---

# Libraries 

```{r setup, include=FALSE}
library("rlang") #toolbox
library("dplyr") #data manipulation
library("here") #toolbox
library("tidyverse") #data manipulation
library("ggpubr") #plotting
library("tidytext") #data manipulation
library("OpenMx") #residualise metrics
library("umx") #residualise metrics
library("knitr") #rmarkdown
library("htmltools") #rmarkdown
library("markdown") #rmarkdown
library("httpuv") #rmarkdown
library("vtable") #summary tables
library("gtsummary") #summary tables
library("NCmisc") #toolbox
library("stringr") #data manipulation
library("rstatix") #effect sizes
```

# Neonatal MRI and infant attachment

## Section I: Prepare dataframe with data and transform potential covariates

Start from spreadsheet with raw times from coding scheme (Williams and Turner, 2020). 
IDs are already filtered in this spreadhseet, based on MRI + attachment criteria. 
Merge attachment, MRI and covariates data in a dataframe.

```{r}
Attachment <- read.csv(here("raw_data", "IDs_WT_rawtimes_filtered.csv"), header=T, dec=".",sep=",")
MRI_structural <- read.csv(here("raw_data", "IDs_MRI_structural.csv"), header=T, dec=".",sep=",")
MRI_DTI <- read.csv(here("raw_data", "IDs_MRI_DTI_gmNODDI.csv"), header=T, dec=".",sep=",")
Covariates <- read.csv(here("raw_data", "IDs_covariates.csv"), header=T, dec=".",sep=",")
Data <- merge(Attachment, MRI_DTI, by="ID")
Data <- merge(Data, MRI_structural, by="ID")
Data <- merge(Data, Covariates, by="ID")
rownames(Data)<-Data$ID
Data$ID <- NULL
```

Avoid scientific notation.

```{r}
options(scipen = 999)
```

Transform potential covariates to categorical nominal (factors without order) or categorical ordinal (factors with order) accordingly.

```{r}
Data$sex = as.factor(Data$sex)
Data$Ethnicity = as.factor(Data$Ethnicity)
Data$Maternal_smoking = as.factor(Data$Maternal_smoking)
Data$Maternal_alcohol = as.factor(Data$Maternal_alcohol)
Data$Maternal_education = factor(Data$Maternal_education, order = TRUE, levels =c("1", "2", "3", "4", "5", "6", "7"))
Data$Maternal_antenatal_anxiety = as.factor(Data$Maternal_antenatal_anxiety)
Data$Maternal_antenatal_depression = as.factor(Data$Maternal_antenatal_depression)
```

## Section II: Demographics of study sample

Check normality of continuous variables to know how to report them (both Shapiro tests and histograms are used to double-check visually).

```{r}
targetdata=Data[,grep('GA_birth|BW|SIMD|age_scan|Maternal_age', colnames(Data))]
apply(targetdata,2,shapiro.test)

for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% hist(main = cols)
}
```

According to above: SIMD, BW, and GA at birth seem non-normally distributed.

Display tables of descriptive statistics: report medians (range) for GA at birth, BW, SIMD and means (SD) for age at scan, according to above.

```{r}

# data wrangling

Data_baseline <- Data[,c(25:37)]

# create table 

table_dem <- tbl_summary(
  Data_baseline,
  by = NULL,
  label = NULL,
  statistic = list(c("GA_birth", "BW", "SIMD") ~ c("{median} ({min}, {max})"),
                   c("age_scan", "Maternal_age") ~ c("{mean} ({sd})"), all_categorical() ~ "{n}"),
  digits = all_continuous()~3,
  type = NULL,
  value = NULL,
  missing = NULL,
  missing_text = NULL,
  sort = NULL,
  percent = NULL,
  include = everything()
)

# export table 

table_dem
table_dem <- as_tibble(table_dem)
colnames(table_dem) <- c("Characteristic", "Study sample (N=133)")
write.csv(table_dem, here("results/demographics", "Study.sample.csv"), row.names=FALSE)

```

## Section III: Demographic comparisons between the included (N=133) and excluded (N=159) sample

```{r}

# show stats and p-values with 3 digits

Pval_fun <- function(x) style_pvalue(x, digits = 3)
Stat_fun <- function(x) style_sigfig(x, digits=3)

```

```{r}

# data reading and wrangling

IDs_all_MRI <- read.csv(here("raw_data", "IDs_all_MRI.csv"), header=T, dec=".",sep=",")
list_to_exclude <- rownames(Data)
Data_excluded <- IDs_all_MRI[ !IDs_all_MRI$ID %in% list_to_exclude, ]
Data_excluded <- as.data.frame(Data_excluded)
colnames(Data_excluded) <- "ID"
Covariates$age_scan <- NULL
Data_excluded <- merge(Data_excluded, Covariates, by="ID")
rownames(Data_excluded) <- Data_excluded$ID
Data_excluded$ID <- NULL
Data_excluded$group <- "excluded"
Data_included <- Data[c(25:29,31:37)]
Data_included$group <- "included"

col_order <- c("GA_birth", "BW", "Prematurity", "sex", "Ethnicity", "Maternal_age", "Maternal_education", "Maternal_smoking", "Maternal_alcohol", "Maternal_antenatal_depression", "Maternal_antenatal_anxiety", "SIMD", "group")

Data_excluded <- Data_excluded[,col_order]
Data_included <- Data_included[,col_order]
Data_dem_comparison <- rbind(Data_excluded, Data_included)

# create table 

table_com <- Data_dem_comparison %>% tbl_summary(by = group,
  label = NULL,
  statistic = list(c("GA_birth", "BW", "SIMD") ~ c("{median} ({min}, {max})"), 
                   c("Maternal_age") ~ c("{mean} ({sd})"), all_categorical() ~ "{n} ({p}%)"),
  digits = everything()~3) %>% 
  add_p(
  test = list(c("GA_birth", "BW", "SIMD") ~ "wilcox.test",
                   c("Maternal_age") ~ "t.test", all_categorical() ~ "chisq.test"),
  pvalue_fun = Pval_fun) %>% 
  separate_p_footnotes() %>% 
  modify_header(
    list(
      statistic ~ "**t-statistic**",
      p.value ~ "**p-value**"
    )
  ) %>%
  modify_fmt_fun(
    list(
      statistic ~ Stat_fun,
      p.value ~ Pval_fun
    ))



# change format and add effect sizes 

table_com <- as_tibble(table_com)
colnames(table_com) <- c("Characteristic", "excluded (N=159)", "study sample (N=133)", "t-statistic", "p-value")
table_com[, 'effect size'] = NA

table_com$"effect size"[which(table_com$Characteristic=="GA_birth")] <- wilcox_effsize(Data_dem_comparison,GA_birth~group)$effsize

table_com$"effect size"[which(table_com$Characteristic=="BW")] <- wilcox_effsize(Data_dem_comparison,BW~group)$effsize

table_com$"effect size"[which(table_com$Characteristic=="Prematurity")] <- cramer_v(Data_dem_comparison$Prematurity, Data_dem_comparison$group)

table_com$"effect size"[which(table_com$Characteristic=="sex")] <- cramer_v(Data_dem_comparison$sex, Data_dem_comparison$group)  

table_com$"effect size"[which(table_com$Characteristic=="Ethnicity")] <- cramer_v(Data_dem_comparison$Ethnicity, Data_dem_comparison$group)

table_com$"effect size"[which(table_com$Characteristic=="Maternal_age")] <- cohens_d(Data_dem_comparison, Maternal_age~group)$effsize

table_com$"effect size"[which(table_com$Characteristic=="Maternal_education")] <- cramer_v(Data_dem_comparison$Maternal_education, Data_dem_comparison$group)  

table_com$"effect size"[which(table_com$Characteristic=="Maternal_smoking")] <- cramer_v(Data_dem_comparison$Maternal_smoking, Data_dem_comparison$group)

table_com$"effect size"[which(table_com$Characteristic=="Maternal_alcohol")] <- cramer_v(Data_dem_comparison$Maternal_alcohol, Data_dem_comparison$group)

table_com$"effect size"[which(table_com$Characteristic=="Maternal_antenatal_depression")] <- cramer_v(Data_dem_comparison$Maternal_antenatal_depression, Data_dem_comparison$group)

table_com$"effect size"[which(table_com$Characteristic=="Maternal_antenatal_anxiety")] <- cramer_v(Data_dem_comparison$Maternal_antenatal_anxiety, Data_dem_comparison$group)

table_com$"effect size"[which(table_com$Characteristic=="SIMD")] <- wilcox_effsize(Data_dem_comparison,SIMD~group)$effsize

# export table
write.csv(table_com, here("results/demographics", "Exc.inc.comparison.csv"), row.names=FALSE)

```

## Section IV: Correlations between attachment behaviours (raw scores)

Check distribution of attachment behaviours

```{r}
targetdata=Data[,grep('_Raw', colnames(Data))]
apply(targetdata,2,shapiro.test)

for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% hist(main = cols)
}


for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% qqnorm(main = cols)
}
```

FNF_Raw seem non-normally distributed.
Run correlations (Spearman tests for FNF_Raw).

```{r}

# spearman correlations

targetdata=Data[,grep("_Raw", colnames(Data))]

cors <- function(df) { 
  # turn all three matrices (r, n, and P into a data frame)
  M <- Hmisc::rcorr(as.matrix(df), type="spearman")
  # return the three data frames in a list return(Mdf)
  Mdf <- map(M, ~data.frame(.x))
}

newdata <- cors(targetdata)

formatted_cors <- function(df){
  cors(df) %>%
    map(~rownames_to_column(.x, var="measure1")) %>%
    map(~pivot_longer(.x, -measure1, names_to="measure2")) %>% 
    bind_rows(.id = "id") %>%
    tidyr::pivot_wider(names_from = id, values_from = value) %>%
    mutate(sig_label = case_when(
      P<0.05 & P>0.01 ~ "*", 
      P<0.01 & P>0.001 ~ "**", 
      P<0.001 ~ "***",
      TRUE ~ " "))}

data_spearman <- formatted_cors(targetdata)
data_spearman = as.data.frame(data_spearman)
colnames(data_spearman) = c("Y", "X", "Estimate", "N", "P_value", "sig_label")
data_FNF <- data_spearman %>% filter(Y == "FNF_Raw")
data_FNF <- data_FNF %>% filter(X == "HD_Raw" | X == "FNF_Raw" |  X == "AA_Raw" )

# pearson correlations

targetdata=Data[,grep("HD_Raw|AA_Raw", colnames(Data))]

cors <- function(df) { 
  # turn all three matrices (r, n, and P into a data frame)
  M <- Hmisc::rcorr(as.matrix(df), type="pearson")
  # return the three data frames in a list return(Mdf)
  Mdf <- map(M, ~data.frame(.x))
}

formatted_cors <- function(df){
  cors(df) %>%
    map(~rownames_to_column(.x, var="measure1")) %>%
    map(~pivot_longer(.x, -measure1, names_to="measure2")) %>% 
    bind_rows(.id = "id") %>%
    pivot_wider(names_from = id, values_from = value) %>%
    mutate(sig_label = case_when(
      P<0.05 & P>0.01 ~ "*", 
      P<0.01 & P>0.001 ~ "**", 
      P<0.001 ~ "***",
      TRUE ~ " "))}

data_pearson <- formatted_cors(targetdata)
data_pearson = as.data.frame(data_pearson)
colnames(data_pearson) = c("Y", "X", "Estimate", "N", "P_value", "sig_label")

# data wrangling 

data_HD <- data_pearson %>% filter(Y == "HD_Raw"| Y == "AA_Raw")
data_HD <- data_HD %>% filter(X == "HD_Raw" )

data_AA <- data_pearson %>% filter(Y == "AA_Raw")
data_AA <- data_AA %>% filter(X == "AA_Raw")

data <- rbind(data_FNF, data_HD, data_AA)

data$X <- factor(data$X, levels=c("HD_Raw", "FNF_Raw", "AA_Raw"))
data$Y <- factor(data$Y, levels=c("AA_Raw", "FNF_Raw", "HD_Raw"))
data <- data %>% mutate_if(is.numeric, ~round(., 2))
data[3,2] <- "FNF_Raw"
data[3,1] <- "AA_Raw"

# create and save heatmap from correlations

g <- ggplot(data = data, aes(x = X, y = Y, fill = Estimate)) +
  scale_fill_gradient2(mid="#FBFEF9",low="#841026",high="#0E6885", limits=c(-1,1)) +
  geom_tile() +
  ylab(label = "Attachment behaviour") +
  xlab(label = "Attachment behaviour") 


g <- g + geom_text(aes(label=Estimate), vjust=0.6, hjust=0.5) + theme_bw()

g

ggsave(here("results/figures", "Beh.cor.heatmap.pdf"), width = 130, height = 100, units = "mm", plot = g, dpi=300)

```

```{r}
markdownToHTML("MRI_Attachment_dem.Rmd",output ="MRI_Attachment_dem.html")
```

